Stochastic Variance-Reduced Policy Gradient

### 2018-02-08 svrpg.tex
expanded section 7

\subsection{Zeroth update}
As noted in Section \ref{sec:alg}, the update performed at the beginning of each epoch is equivalent to a full-gradient update. In our setting, where collecting samples is particularly expensive, the $B$ trajectories collected using the snapshot trajectory $\pi_{\tilde{\vtheta}^s}$ feel like a waste of data. In practice, we just perform an approximate full gradient update using the $N$ trajectories sampled to compute $\gradApp{\tilde{\vtheta}^s}{N}$, \ie
\begin{align*}
	&\vtheta_{1}^{s+1} = \tilde{\vtheta}^s + \alpha\gradApp{\tilde{\vtheta}^s}{N} \\
	&\vtheta_{t+1}^{s+1} = \vtheta_t^{s+1} + \alpha v_t^{s+1} \text{ for $t=1,\dots,m-1$}.
\end{align*}
In the rest of this section we assume this practical variant.

@@

\section{Experiments}\label{sec:exp}
In this section we address the performance evaluation of our policy gradient estimator defined in equation \ref{E:svrpg.estimate.batch} against that one of equation \ref{E:policygradient.estimate}. This comparison is performed using G(PO)MDP as estimator of $\score{\vtheta}{\tau}\Reward(\tau)$. For the sake of simplicity and with a little abuse of notation we refer to \ref{E:svrpg.estimate.batch} as SVRPG and to \ref{E:policygradient.estimate} as G(PO)MDP.
In all the experiments we used gaussian policies and bootsrap confidence intervals $\alpha = 10\% $ over ten different runs for evaluation. Each run with a different policy initialization which is kept fixed when changing the algorithm. The maximum number of trajectories at hand is fixed for each task. In \emph{figure} \ref{fig:swimmerone} we can see how SVRPG outperforms G(PO)MDP over the swimmer task. Clearly the mean is nearly always higher for our algorithm, moreover near the end we see that te confidence intervals are also disjoint. In \emph{figure} \ref{fig:swimmertwo} we addressed the comparison between SVRPG and its self-normalized version still for the swimmer task. We can see an inprovement on the average return, but the confidence intervals are never disjoint. In \emph{figure} \ref{fig:halfcheetah} the difference between our algorithm and G(PO)MDP over the half-cheetah task is not statistically significant (both of them use the same baseline even though is not optimal for SVRPG see \ref{} for further hints over this topic), however we can see a certain improvement on the mean of the objective function. Eventually in \emph{figure} \ref{fig:cartpole} we can clearly see that SVRPG outperforms GPOMDP in the cart-pole task.

\begin{figure}[t]
    \includegraphics[width=0.5\textwidth]{swimmer_self_normalized_SVRPG_vs_GPOMDP}
    \caption{Self-Normalized SVRPG vs GPOMDP in swimmer}
    \label{fig:swimmerone}
\end{figure}
\begin{figure}[t]
    \includegraphics[width=0.5\textwidth]{swimmer_self_normalized_SVRPG_vs_SVRPG}
    \caption{Self-Normalized SVRPG vs Non Self-Normalized SVRPG in swimmer}
    \label{fig:swimmertwo}
\end{figure}
\begin{figure}[t]
    \includegraphics[width=0.5\textwidth]{half_cheetah_self_norm_baseline_SVRPG_vs_GPOMDP100_baseline}
    \caption{Self-Normalized SVRPG with baseline vs GPOMDP with baseline in half cheetah}
    \label{fig:halfcheetah}
\end{figure}
\begin{figure}[t]
    \includegraphics[width=0.5\textwidth]{cart_pole_SVRPG_vs_GPOMDP}
    \caption{SVRPG vs GPOMDP in cart-pole}
    \label{fig:cartpole}
\end{figure}

\section{Conclusion}
In this paper, we introduced SVRPG, a variant of SVRG designed explicitly for RL problems.
The control problem considered in the paper has a series of difficulties that are not common in SL.
Among them, non-concavity and approximate estimates of the FG have been analysed independently in SL~\citep[\eg][]{allen2016variance,reddi2016stochastic,harikandeh2015stopwasting} but never combined.
Nevertheless, the main issue in RL is the non-stationarity of the sampling process since the distribution underlying the objective function is policy-dependent.
We have shown that by exploiting importance weighting techniques, it is possible to overcome this issue and preserve the unbiasedness of the corrected gradient.
We have additionally shown that, under mild assumptions that are often verified in RL applications, it is possible to derive convergence guarantees for SVRPG.
{\color{red}SVRG has been previously used in RL for policy evaluation~\citep{du2017svrgpe} and as alternative optimization method for TRPO~\citep{xu2017svrgtrpo}.
The first case is much more similar to the SL case since the problem is stationary and may be made concave, removing several issues we have faced here.
In the second case, it was used as a better optimization technique to solve the per-iteration problem imposed by TRPO.
}
\todomat{Repetition of Section 6?}
Finally, we have empirically shown that practical variants of the theoretical SVRPG version can outperform classical actor-only approaches on several benchmarks.

>>>>

\subsection{Zeroth update}
As noted in Section \ref{sec:alg}, the update performed at the beginning of each epoch is equivalent to a full-gradient update. In our setting, where collecting samples is particularly expensive, the $B$ trajectories collected using the snapshot trajectory $\pi_{\tilde{\vtheta}^s}$ feel like a waste of data. In practice, we just perform an approximate full gradient update using the $N$ trajectories sampled to compute $\gradApp{\tilde{\vtheta}^s}{N}$, \ie
\begin{align*}
	&\vtheta_{1}^{s+1} = \tilde{\vtheta}^s + \alpha\gradApp{\tilde{\vtheta}^s}{N} \\
	&\vtheta_{t+1}^{s+1} = \vtheta_t^{s+1} + \alpha v_t^{s+1} \text{ for $t=1,\dots,m-1$}.
\end{align*}
In the following we will always assume this practical variant is employed.

@@

\section{Experiments}\label{sec:exp}
In this section we evaluate the performance of SVRPG and compare it with policy gradient (PG) on well known continuous RL tasks. This comparison is performed using G(PO)MDP as an estimator of $\score{\vtheta}{\tau}\Reward(\tau)$ in both cases. For our algorithm, we use a batch size $N=100$, a mini-batch size $B=10$, and the jointly adaptive step size $\alpha$ and epoch length $m$ proposed in Section \ref{sec:stopping}. Since the aim of this comparison is to show the improvement that SVRG-flavored variance reduction brings to SG in the policy gradient framework, we set the batch size of the baseline policy gradient algorithm to $B$. In this sense, we measure the improvement yielded by computing snapshot gradients and using them to adjust parameter updates. Since we evaluate on-line performance over the number of sampled trajectories, the cost of computing such snapshot gradients is automatically taken into consideration. To make the comparison fair, we also use Adam in the baseline PG algorithm, which we will denote simply as G(PO)MDP in the following.
In all the experiments we used deep Gaussian policies with adaptive standard deviation (details on network architecture in Appendix \ref{app:exp}). Each experiment is run 10 times with a random policy initialization, different each time, shared by all the algorithms under comparison. The length of the experiment, \ie the total number of trajectories, is fixed for each task. Performance is evaluated by using test-trajectories on a subset of the policies considered during the learning process. Of course test trajectories are not counted, since they are only meant to improve the accuracy of reported results and do not affect learning. We provide average performance with 90\% bootstrap confidence intervals.  
Task implementations are from the \textit{rllab} library \cite{duan2016benchmarking}\footnote{https://github.com/rll/rllab}, on which our agents are also based.

\paragraph{Cart-pole.} Figure \ref{fig:cartpole} compares SVRPG with G(PO)MDP on the continuous cart-pole balancing task. Here the performance gain is evident. Notice also how the variance of our algorithm is visibly smaller.

\begin{figure}[t]
	\includegraphics[width=0.5\textwidth]{cart_pole_SVRPG_vs_GPOMDP}
	\caption{SVRPG vs G(PO)MDP in the Cart-pole task}
	\label{fig:cartpole}
\end{figure}

\paragraph{Swimmer} The mujoco swimmer task is far more difficult than cart-pole. The longer horizon and the more complex dynamics can have a dangerous impact on the variance of importance weights. In this case, the self-normalization technique proposed in Section \ref{sec:selfnorm} brings an improvement, as shown in Figure \ref{fig:swimmertwo}. Although the confidence intervals are never disjoint, the average of the self-normalized version is always higher, which we consider enough to employ the variant in the subsequent comparison. Figure \ref{fig:swimmerone} shows self-normalized SVRPG against G(PO)MDP. On average, our algorithm dominates for almost the entire task. Toward the end of the task, the improvement becomes statistically significant. Also in this case, the different width of confidence intervals suggest that SVRPG is more robust, especially in the early iterations.

\begin{figure}[t]
	\includegraphics[width=0.5\textwidth]{swimmer_self_normalized_SVRPG_vs_SVRPG}
	\caption{Self-Normalized SVRPG vs unbiased SVRPG in the Swimmer task}
	\label{fig:swimmertwo}
\end{figure}

\begin{figure}[t]
	\includegraphics[width=0.5\textwidth]{swimmer_self_normalized_SVRPG_vs_GPOMDP}
	\caption{Self-Normalized SVRPG vs G(PO)MDP in the Swimmer task.}
	\label{fig:swimmerone}
\end{figure}

%\begin{figure}[t]
%    \includegraphics[width=0.5\textwidth]{half_cheetah_self_norm_baseline_SVRPG_vs_GPOMDP100_baseline}
%    \caption{Self-Normalized SVRPG with baseline vs GPOMDP with baseline in half cheetah}
%    \label{fig:halfcheetah}
%\end{figure}

\section{Conclusion}
In this paper, we introduced SVRPG, a variant of SVRG designed explicitly for RL problems.
The control problem considered in the paper has a series of difficulties that are not common in SL.
Among them, non-concavity and approximate estimates of the FG have been analysed independently in SL~\citep[\eg][]{allen2016variance,reddi2016stochastic,harikandeh2015stopwasting} but never combined.
Nevertheless, the main issue in RL is the non-stationarity of the sampling process since the distribution underlying the objective function is policy-dependent.
We have shown that by exploiting importance weighting techniques, it is possible to overcome this issue and preserve the unbiasedness of the corrected gradient.
We have additionally shown that, under mild assumptions that are often verified in RL applications, it is possible to derive convergence guarantees for SVRPG.
{\color{red}SVRG has been previously used in RL for policy evaluation~\citep{du2017svrgpe} and as alternative optimization method for TRPO~\citep{xu2017svrgtrpo}.
The first case is much more similar to the SL case since the problem is stationary and may be made concave, removing several issues we have faced here.
In the second case, it was used as a better optimization technique to solve the per-iteration problem imposed by TRPO.
}
\todomat{Repetition of Section 6?}
Finally, we have empirically shown that practical variants of the theoretical SVRPG version can outperform classical actor-only approaches on benchmark tasks.
